Bringing the Power of eBPF to Open vSwitch

### 2018-10-22 paper.tex
revise introducation

\begin{abstract}
eBPF is an emerging technology in Linux kernel with the goal of making Linux kernel
extensible by providing an eBPF virtual machine with safety guarantee.
Open vSwitch, OVS, is a software switch running majorly in Linux
operating systems and interacts with Linux kernel module,
the openvswitch.ko.
In this paper we describe our efforts to make use of eBPF technology in OVS.
We present the design, implementation, and evaluation of two projects:
the OVS-eBPF project and the OVS-AFXDP project.
OVS-eBPF has the goal of re-writing existing flow processing features
in openvswitch kernel module into eBPF code.
On the other hand, the OVS-AFXDP project aims to by-passing the kernel
using AF\_XDP socket and moving most of the flow processing features
in userspace.

@@

eBPF, extended Berkeley Packet Filter, enables userspace applications to
customize and extend the Linux kernel's functionality.  It provides
flexible platform abstractions for network functions, and is being ported
to a variety of platforms. 
The Linux kernel BPF~\cite{cbpf,bpfp},
with its extension called eBPF~\cite{lwnebpf,daniel1,daniel2},
provides the desired infrastructure for an OVS datapath implementation.
Specifically, a packet flowing inside a Linux
box can be intercepted and trigger the execution of arbitrary eBPF programs.

Among the various ways of using eBPF, OVS has been exploring the power of eBPF in three:
(1) attaching eBPF to TC, (2) offloading a subset of processing to XDP, and (3) by-passing
the kernel using AF\_XDP.
Attaching eBPF to TC started first with the most aggressive goal: we planned to re-implement
the entire features of OVS kernel datapath under net/openvswitch/* into eBPF code.
We worked around a couple of limitations, for example, the lack of TLV support led us to
redefine a binary kernel-user API using a fixed-length array; and without a dedicated way to
execute a packet, we created a dedicated device for user to kernel packet transmission,
with a different BPF program attached to handle packet execute logic.
Currently, we are working on connection tracking.
Although a simple eBPF map can achieve basic operations of conntrack table lookup and commit,
how to handle NAT, (de)fragmentation, and ALG are still under discussion.

Moving one layer below TC is called XDP (eXpress Data Path), a much faster layer for packet
processing, but with almost no extra packet metadata and limited BPF helpers support.
Depending on the complexity of flows, OVS can offload a subset of its flow processing to XDP
when feasible. However, the fact that XDP has fewer helper function support implies that either
1) only very limited number of flows are eligible for offload, or 2) more flow processing logic
needed to be done in native eBPF.

AF\_XDP is another form of XDP but providing a socket interface for control plane and a shared
memory API for accessing packets from userspace application. OVS today has another full-fledged
datapath implementation in userspace, called dpif-netdev, used by DPDK community.
By treating the AF\_XDP as a fast packet-I/O channel, the OVS dpif-netdev can satisfy most of the
features. We are working on building the prototype and evaluating its performance.

Existing OVS consists of three datapath implementations:
dpif-netlink
dpif-netdev
dpif-windows
In this paper, OVS-eBPF project adds a new datapath type
dpif-bpf
and OVS-AFXDP project reuses the dpif-netdev interface, with
linux netdev receiving/sending packets from AFXDP sockets.

Table~\ref{compare} shows the difference between two projects:
\begin{table}
\centering
\scriptsize
\begin{tabular}{|c | c | c|}
\hline
 {\bf Projects} & {\bf OVS-eBPF } & {\bf OVS-AFXDP} \\ \hline\hline
%  Performance & $<$2Mpps & $>14$Mpps \\ \hline
  Performance & Low & High \\ \hline
  Driver Support & No & Yes \\ \hline
  Minimal Kernel & 4.18 & Unknown \\ \hline
  BPF Code Size & Large & Minimal \\ \hline
  Extensibility & Low & High \\ \hline
  \end{tabular}
\caption{\footnotesize
Comparison of the OVS-eBPF project and OVS-AFXDP project.
}
\label{compare}
\end{table}

@@

\mycomment{
As each new feature in OVS datapath requires deep knowledge and programming
paradigm of the target operating system, e.g., how Linux packets are received,
how Windows kernel module does firewalling, and each operating system
community has its own software upstream and maintenance processes, the software
development cycle of a new datapath feature becomes complex and requires
detailed platform-specific knowledge. Each platform introduces its own
development and maintenance burden, and it is common for new features to take
six months or more before implementations are available across all datapaths.
As a result, providing a programmable OVS datapath which allows extending the
features flexibly with lower maintenance cost becomes a critical requirement.
}

@@

\subsubsection{Memory Management and Locks}
For every stage of the optimization, we use Linux perf extensively, e.g., perf stat
and perf record/report.  With the above new design, we observe the new bottleneck
is at the UMEM memory pool we introduce to maintain the UMEM freelist, and the
packet metadata allocation, struct dp\_packet, for every receiving packet.
We introduce two major API, umem\_elem\_get(), which gets N free element from the
UMEM freelist, and umem\_elem\_put(), which places back the free UMEM buffer to
the freelist.
producer 
consumer
We implemented three data structures of UMEM memory pool as below
to compare their performance
\begin{itemize}
\item list\_head: Similar to Linux kernel's list\_head, each element contains it
UMEM addr, and a next pointer points to next free element. Elements are sparse in
memory. Consumer always takes the first element, head, and producer also place back
from the head of the list.
\item ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
allocated in continuous memory region, and pointers (head and tail) point for
Consumer takes elements from the tail and when placing back, producer, puts
them into where head pointer points to.
\item ptr\_stack: Similar to ptr\_ring, an array of elements are allocated
in continuous memory region, but both consumer and producer takes elements
from beginning of the array.
\end{itemize}

@@

%\subsubsection{Metadat Allocation}
With the above change, Linux perf shows that the packets metadata allocation
takes a lot of CPU cycles, the dp\_packet\_init, dp\_packet\_set\_data.
So instead of allocating packet metadata at packet reception time, we
implement two data structures and compare their performance:
\begin{itemize}
\item Inlining in UMEM buffer: Since we already allocate 2K chunk for each
UMEM buffer, we reserve the first 256-byte in each UMEM element as struct dp\_packet
and initialize the dp\_packet as much as we can at allocation time.
So when packets arrive, each packet already has its packet metadata at a const
offset ahead.
\item Array on heap: This design simply allocate a continuous memory region
storing an array of packet metadata, and initialize them as much as we can.
\end{itemize}

%complete separation of the allocation of metadata structures from the allocation of packet data buffers.
% see mbuf dpdk https://doc.dpdk.org/guides/prog_guide/mbuf_lib.html
%Embed metadata within a single memory buffer the structure followed by a fixed size area for the packet data.
%Use separate memory buffers for the metadata structure and for the packet data.

>>>>

\begin{abstract}
eBPF is an emerging technology in Linux kernel with the goal of making Linux kernel
extensible by providing an eBPF virtual machine with safety guarantee.
Open vSwitch, OVS, is a software switch running majorly in Linux
operating systems and interacts with Linux kernel module,
the openvswitch.ko.
In this paper we describe our efforts to make use of eBPF technology in OVS.
We present the design, implementation, and evaluation of two projects:
the OVS-eBPF project and the OVS-AFXDP project.
OVS-eBPF has the goal of re-writing existing flow processing features
in openvswitch kernel module into eBPF code, by attaching eBPF program
to Linux TC.
On the other hand, the OVS-AFXDP project aims to by-passing the kernel
using AF\_XDP socket and moves most of the flow processing features
into userspace.

@@

eBPF, extended Berkeley Packet Filter, enables userspace applications to
customize and extend the Linux kernel's functionality.  It provides
flexible platform abstractions for network functions, and is being ported
to a variety of platforms.  Among the various different ways of using eBPF,
OVS has been exploring the power of eBPF in three:
1) in-kernel flow processing by attaching eBPF programs to TC,
2) offloading a subset of processing to XDP (eXpress Data Path),
and 3) moving the flow processing to userspace by using AF\_XDP.

Attaching eBPF to TC, OVS-eBPF project, started first with the most aggressive goal:
we planned to re-implement the entire features of OVS kernel datapath under net/openvswitch/* into eBPF code.
We worked around a couple of limitations, for example, the lack of TLV (Type-Length-Value) support
led us to redefine a binary kernel-user API using a fixed-length array;
and without a dedicated way to execute a packet, we created a dedicated
device for user to kernel packet transmission,
with a different BPF program attached to handle packet execute logic.
Currently, OVS-eBPF can satisfy most of the basic feature requirement
for flow processing, including tunneling protocol support.
But more complicated processing such as connection tracking, NAT,
(de)fragmentation, and ALG are still under discussion.

Moving one layer below TC is called XDP (eXpress Data Path), a much faster layer for packet
processing, but with almost no extra packet metadata and limited BPF helpers support.
Depending on the complexity of flows, OVS can offload a subset of its flow processing to XDP
when feasible. However, the fact that XDP has fewer helper function support implies that either
1) only very limited number of flows are eligible for offload, or 2) more flow processing logic
needed to be done in native eBPF code.

XDP provides another way for interacting with userspace programs, called AF\_XDP. 
AF\_XDP is another socket interface for control plane and a shared
memory API for accessing packets from userspace application.
OVS today has another full-fledged
datapath implementation in userspace, called dpif-netdev, mostly used by DPDK community.
OVS-AFXDP project treats the AF\_XDP as a fast packet-I/O channel, and move
most of the flow processing into the userspace OVS.

A rough comparison of the two project is shown in Table~\ref{compare}
This paper describes the design, implementation, and evaluation of the
OVS-eBPF and OVS-AFXDP projects.
\begin{table}
\centering
\scriptsize
\begin{tabular}{|c | c | c|}
\hline
 {\bf Projects} & {\bf OVS-eBPF } & {\bf OVS-AFXDP} \\ \hline\hline
  Packet Rate & Low & High \\ \hline
  Driver Support & No & Yes \\ \hline
  Minimal Kernel Req. & 4.18 & Unknown \\ \hline
  BPF Code Size & Large & Minimal \\ \hline
  Extensibility & Low & High \\ \hline
  \end{tabular}
\caption{\footnotesize
Comparison of the OVS-eBPF project and OVS-AFXDP project.
}
\label{compare}
\end{table}

@@

@@

\subsubsection{Memory Management and Locks}
For every stage of the optimization, we use Linux perf extensively, e.g., perf stat
and perf record/report.  With the above new design, we observe the new bottleneck
is at the UMEM memory pool we introduce to maintain the UMEM freelist, and the
packet metadata allocation, struct dp\_packet, for every receiving packet.
We introduce two major API, umem\_elem\_get(), which gets N free element from the
UMEM freelist, and umem\_elem\_put(), which places back the free UMEM buffer to
the freelist.
producer, consumer
We implemented three data structures of UMEM memory pool as below
to compare their performance
\begin{itemize}
\item list\_head: Similar to Linux kernel's list\_head, each element contains it
UMEM addr, and a next pointer points to next free element. Elements are sparse in
memory. Consumer always takes the first element, head, and producer also place back
from the head of the list.
\item ptr\_ring: Similar to Linux kernel's ptr\_ring, an array of elements are
allocated in continuous memory region, and pointers (head and tail) point for
Consumer takes elements from the tail and when placing back, producer, puts
them into where head pointer points to.
\item ptr\_stack: Similar to ptr\_ring, an array of elements are allocated
in continuous memory region, but both consumer and producer takes elements
from beginning of the array.
\end{itemize}

@@

\subsubsection{Metadat Allocation}
With the above change, Linux perf shows that the packets metadata allocation
takes a lot of CPU cycles, the dp\_packet\_init, dp\_packet\_set\_data.
So instead of allocating packet metadata at packet reception time, we
implement two data structures and compare their performance:
\begin{itemize}
\item Embedding in UMEM buffer: Since we already allocate 2K chunk for each
UMEM buffer, we reserve the first 256-byte in each UMEM element as struct dp\_packet
and initialize the dp\_packet as much as we can at allocation time.
This is similar to the DPDK mbuf design~\cite{dpdkmbuf}, where a single memory buffer
is used for both packet data and metadata.
\item Array on heap: This design simply allocate a continuous memory region
storing an array of packet metadata, and initialize them as much as we can.
\end{itemize}

%complete separation of the allocation of metadata structures from the allocation of packet data buffers.
% see mbuf dpdk https://doc.dpdk.org/guides/prog_guide/mbuf_lib.html
%Embed metadata within a single memory buffer the structure followed by a fixed size area for
% the packet data.
%Use separate memory buffers for the metadata structure and for the packet data.
